---
title: "Introduction to package"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(shiny)
library(clinicalfd)

library (tidyverse)
library(learnr)
library(gt)
knitr::opts_chunk$set(echo = FALSE)
```
## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
adsl <- clinicalfd::adsl
head(adsl)
```

 

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

```{r quiz }
quiz(
  question("How do read clinicalfd package into Rstudio?",
    answer("Just type in Clinicalfd"),
    answer("just type in clinicalfd"),
    answer("Just type in install(clinicalfd) ", correct = TRUE),
    answer("Do not type anything")
  ),
  
  question("How to use ADSL dataset and usubjid ?",
    answer("adsl <- clinicalfd::adsl ", correct = TRUE),
    answer("ADSL"),
    answer("clinicalfd.adsl"),
    answer("select (clinical::adsl , usubjid) ", correct = TRUE)
  )
)
```

## GT introduction

- All **gt** tables begin with a call to `gt()`, where we would supply     the input data table (data frame or tibble) and some basic options for   creating a stub (`rowname_col`) and row groups (`groupname_col`).
- We can also introduce a grouped tibble (via **dplyr**’s `group_by()`)    for more precise divisions of rows into row groups. 
- The `gt_preview()` function is great for getting a nicely-formatted preview of a data table (defaulting to the first 5 rows and the last row).
- Finally, sometimes we want to export the table to some format. It could be HTML, it could be RTF... whichever you go with, you can use the `gtsave()` function.

### Functions in this module 
- 'gt()'
- 'gt_preview()'
- 'gt_save()'
 
 
## `gt()`: Create a **gt** table object

``` r
gt(
  data,
  rowname_col = "rowname",
  groupname_col = dplyr::group_vars(data),
  rownames_to_stub = FALSE,
  auto_align = TRUE,
  id = NULL,
  row_group.sep = getOption("gt.row_group.sep", " - ")
)
```
### Get Help using ?gt

- The gt() function creates gt() table object when provided with table data.
- Using this function is the first step in typical **gt** workflow. 
- Once we have the **gt** table object , we can perform styling transformations rendering to a display table of various formats

##### EXAMPLES

The **gt** package contains a few datasets:

- `countrypops` - Yearly populations of countries from 1960 to 2017
- `sza` - Twice hourly solar zenith angles by month & latitude
- `gtcars` - Deluxe automobiles from the 2014-2017 period
- `sp500` - Daily S&P 500 Index data from 1950 to 2015
- `pizzaplace` - A year of pizza sales from a pizza place
- `exibble` - A toy example tibble for testing with **gt**: `exibble`

Let's print out `exibble` and see what it looks like:

## Now let's introduce dataset to gt()

```{r exibble , echo = TRUE}
 exibble %>% gt(rowname_col = "row")
```


