---
title: "Introduction to package"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(shiny)
library(clinicalfd)

library (tidyverse)
library(learnr)
library(gt)
knitr::opts_chunk$set(echo = FALSE)
```
## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
adsl <- clinicalfd::adsl
head(adsl)
```

```{r}
# slice_sample() randomly selects rows.
adsl02 <- adsl %>% 
           slice_sample( n = 30) %>% 
            select ( subjid , trt01a , age , sex ) 
 

```


```{r  adsl , echo = TRUE}
 adsl02  %>% gt(rowname_col = "trt01a")
```


 

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

```{r quiz }
quiz(
  question("How do read clinicalfd package into Rstudio?",
    answer("Just type in Clinicalfd"),
    answer("just type in clinicalfd"),
    answer("Just type in install(clinicalfd) ", correct = TRUE),
    answer("Do not type anything")
  ),
  
  question("How to use ADSL dataset and usubjid ?",
    answer("adsl <- clinicalfd::adsl ", correct = TRUE),
    answer("ADSL"),
    answer("clinicalfd.adsl"),
    answer("select (clinical::adsl , usubjid) ", correct = TRUE)
  )
)
```

## GT introduction

- All **gt** tables begin with a call to `gt()`, where we would supply     the input data table (data frame or tibble) and some basic options for   creating a stub (`rowname_col`) and row groups (`groupname_col`).
- We can also introduce a grouped tibble (via **dplyr**’s `group_by()`)    for more precise divisions of rows into row groups. 
- The `gt_preview()` function is great for getting a nicely-formatted preview of a data table (defaulting to the first 5 rows and the last row).
- Finally, sometimes we want to export the table to some format. It could be HTML, it could be RTF... whichever you go with, you can use the `gtsave()` function.

### Functions in this module 
- 'gt()'
- 'gt_preview()'
- 'gt_save()'
 
 
## `gt()`: Create a **gt** table object

``` r
gt(
  data,
  rowname_col = "rowname",
  groupname_col = dplyr::group_vars(data),
  rownames_to_stub = FALSE,
  auto_align = TRUE,
  id = NULL,
  row_group.sep = getOption("gt.row_group.sep", " - ")
)
```
### Get Help using ?gt

- The gt() function creates gt() table object when provided with table data.
- Using this function is the first step in typical **gt** workflow. 
- Once we have the **gt** table object , we can perform styling transformations rendering to a display table of various formats

##### EXAMPLES

The **gt** package contains a few datasets:

- `countrypops` - Yearly populations of countries from 1960 to 2017
- `sza` - Twice hourly solar zenith angles by month & latitude
- `gtcars` - Deluxe automobiles from the 2014-2017 period
- `sp500` - Daily S&P 500 Index data from 1950 to 2015
- `pizzaplace` - A year of pizza sales from a pizza place
- `exibble` - A toy example tibble for testing with **gt**: `exibble`

Let's print out `exibble` and see what it looks like:

## Now let's introduce dataset to gt()

```{r exi , echo= TRUE}
exibble
```

- The above is basic layout , containing the column labels and body cells
- Numbers are minimally formatted 
- We can modify the layout little bit in the first step and create stub. A
  is a special column that contains row labels/names. 

- To do this we supply the name of the column we'd like to use as the stub
to *row_column* argument .Conveniently we have the **row** column in exibble that's good purpose of this 


```{r exibble , echo = TRUE}
 exibble %>% gt(rowname_col = "row")
```
## Gt() tab_stubhead()
Great! A bit on the stub. It doesn’t have a column label above it because it’s technically not a column (it’s the… stub). We can still put a label above it though! That’s with the tab_stubhead() function (in the next module).

We can do some more structuring. Sets of rows can be grouped, showing a row group label above each group. We can do this by using a column containing categorical values (usually grouping labels). Conveniently again, exibble has the "group" column. Let’s use that in the rowname_col argument of gt() and see what it looks like:

```{r}
exibble %>% gt(rowname_col = "row" , groupname_col = "group")
```

Even if rows are in a weird order, gt will put the rows in the correct group (but still respect the order of rows).

```{r}
exibble %>%  dplyr::sample_n( size= 8) %>%  gt(rowname_col = "row" 
                                               , groupname_col = "group")
```

```{r}
mtcars
```
### rownames_to_stub get lines under each row and prints in a nice format
```{r}
 mtcars %>%   gt(rownames_to_stub = TRUE)
```
### using *dplyr group_by * -ing . if you have a grouped tibble , you can pass that into gt() and it will know what to do 

```{r}
# what this does is that use dplyr group_by  to group the rows and print by group 
exibble %>%  dplyr:: group_by(group) %>%  gt(rowname_col = "row")

# very similar to 

adsl02 %>% dplyr:: group_by(sex , age )  %>%  gt(rowname_col = 'trt01' )
```

### This also works when they are multiple columns grpuped .Use **row_group.sep** to specify separator between the labels 

```{r}
pizzaplace %>% 
  dplyr:: filter (type %in% c ("supreme" ,"veggie") ) %>% 
  dplyr:: group_by( type , size) %>% 
  dplyr:: slice_head( n= 3 ) %>% 
  gt(row_group.sep =  "  // ")

# NOTE dplyr:: can be used with every dplyr function or just once
#gt(row_group.sep - seprates rows after grouping them ) printing under the subjid
adsl02 %>%  
  dplyr:: filter (age > 40 ) %>% 
  group_by (age , sex ) %>% 
  slice_head( n = 5 ) %>% 
  gt (row_group.sep =  " <------>"  )

#Note row_group.sep does not work with , rowname_col = "Age & Sex" and , groupname_col = "Group"
  
```

### *gt_previw() preview the a gt table object *
-Sometimes you may want to see just a small portion of your input data. 
We can use gt_preview() in place of gt() to get the first x rows of data and the last y rows of data
(which can be set by the top_n and bottom_n arguments).

-It's not advised to use additional gt functions to further modify the output of gt_preview().
-Furthermore, you cannot pass a gt object to gt_preview().

-           gt_preview(data,
            top_n = 5,
            bottom_n = 1,
            incl_rownums = TRUE)
  

```{r}
#names(adsl)

adsl  %>%  
  dplyr:: select ( subjid , age , sex , ethnic , arm ) %>% 
  gt_preview(  bottom_n = 3 )
  

# TO exclude the line numbers
adsl  %>%  
  dplyr:: select ( subjid , age , sex , ethnic , arm ) %>% 
  gt_preview( top_n = 3 , bottom_n = 3  , incl_rownums = FALSE)
```

##  gtsave() : **Save a gt table to a file**

```{r}
#?gtsave()
getwd()
gt_tbl <- 
  exibble %>% 
   gt(rowname_col = "row" , 
      groupname_col =  "group")
gt_tbl 
```


###  Use **'devtools::install_github("rstudio/gt") '** while installing this package lot of dependencies are involved , if the rlnguage() fails to install try installing tidyverse package then see if it works . If that is not working try deleting .000lock file int the package folder  **'C:\Users\tatav\OneDrive\Documents\R\win-library\3.6'** 
### Installation takes lot of time ( atleast 30 min )


```{r}
 gt_tbl %>%  gtsave ("tab_r.rtf")

```


```{r}
tab_rtf <- gtcars %>% 
           dplyr:: select (mfr , model) %>% 
           dplyr:: slice (1:2) %>% 
           gt() %>% 
           tab_header (
             title = md ("Data listing from **gtcars** "), 
             subtitle = md ("gtcars is an R dataset")
             )%>% 
            as_rtf()
           
    
```


## Adding or Modifying Parts of a Table

A **gt** table can contain a few useful parts for conveying additional information. These include a header (with a titles and subtitle), a footer (with footnotes and source notes), and additional areas for labels (row group labels, column spanner labels, the stubhead label). We can modify the look of table parts more generally with `tab_options()` and perform styling on targeted table locations with `tab_style()`.

------

Functions in this module

- `tab_stubhead()`
- `tab_spanner()`
- `tab_header()`
- `tab_source_note()`
- `tab_footnote()`
- `tab_style()`
- `tab_options()`

 

Helpers for selecting columns:

- `vars()`
- `contains()`
- `matches()`
- `starts_with()`
- `ends_with()`
- `everything()`
 

Helpers for transforming text:

- `md()`
- `html()`
 

Helpers for targeting locations:

- `cells_title()`
- `cells_stubhead()`
- `cells_column_spanners()`
- `cells_column_labels()`
- `cells_row_groups()`
- `cells_stub()`
- `cells_body()`
- `cells_summary()`
- `cells_grand_summary()`


Helpers for defining styles:

- `cell_text()`
- `cell_fill()`
- `cell_borders()`

------



##  `tab_stubhead()`: Add label text to the stubhead

``` r
tab_stubhead(
  data,
  label
)
```
 
Add a label to the stubhead of a **gt** table. The stubhead is the lone element that is positioned left of the column labels, and above the stub. We have the flexibility to use Markdown formatting for the stubhead label with `md()`. Furthermore, if the table is intended for HTML output, we can use HTML for the stubhead label (with `html()`).

##### EXAMPLE

Use `gtcars` to create a **gt** table. Add a stubhead label to describe what is in the stub.

```{r}
#gtcars
```


```{r }
gtcars %>%
  dplyr::select(model, year, hp, trq) %>%
  dplyr::slice(1:5) %>%
  gt(rowname_col = "model") %>%
  tab_stubhead(label = "car")
```


### `tab_spanner()`: Add a spanner column label

``` r
tab_spanner(
  data,
  label,
  columns,
  gather = TRUE
)
```

Set a spanner column label by mapping it to columns already in the table. This label is placed above one or more column labels, spanning the width of those columns and column labels.

With `columns` we can use column names in double quotes (`"<column>"`), in `vars()` (`vars(<column>)`), or, we can use the following **tidyselect** expressions:

- `contains()`: contains a literal string
- `matches()`: matches a regular expression
- `starts_with()`: starts with a prefix
- `ends_with()`: ends with a suffix
- `everything()`: selects all columns

##### EXAMPLE

Let's use the `gtcars` table, but cut it down to size first:

```{r}
gtcars_small <- gtcars %>%  
        dplyr:: select ( model, year , bdy_style, hp , hp_rpm , trq ,
                         trq_rpm , mpg_c , mpg_h , msrp ) %>% 
  dplyr:: slice (1: 8 )

gtcars_small
```

### Use *"gtcars"* to create *"gt"* table ; Group several columns related to car performance under a spanner column with the label *'performance'* 

- *'tab_spanner'* would hlep us create listing style template where create a line under relating treatment arms . 
 - *'label'* is the label used for those columns 
 - *'vars'* is the list of the variables 

```{r}

gtcars_small %>%  
  gt( rowname_col = "model" )  %>% 
  tab_spanner (
  label = "Performance" , 
  column = vars(hp , hp_rpm , trq , trq_rpm , mpg_c , mpg_h)
                    )  
                  
```



